---
title: "Classification Homework"
author: "Tao He"
date: "1/30/2022"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE,out.width="0.9\\linewidth", fig.align  = 'center')
pacman::p_load(
ggplot2,
rstan,
rstanarm,
dplyr,
magick,
ISLR2,
MASS,
caTools,
class,
e1071
)
```

4.6.Suppose we collect data for a group of students in a statistics class with variables X1 = hours studied, X2 = undergrad GPA, and Y = receive an A. We fit a logistic regression and produce estimated coefficient, $$\hat\beta_0$$ = −6, $$\hat\beta_1$$ = 0.05, $$\hat\beta_2$$ = 1.

(a) Estimate the probability that a student who studies for 40 h and has an undergrad GPA of 3.5 gets an A in the class.

***sol.n:***
The logistic function is 

$$P(X) = \frac{\exp(\hat\beta_0 + \hat\beta_1X_1 + \hat\beta_2X_2)}{1 + \exp(\hat\beta_0 + \hat\beta_1X_1 + \hat\beta_2X_2)}$$

Then, let X1 = 40, X2 = 3.5,

$$P(X) = \frac{\exp(-6 + 0.05 * X_1 + X_2)}{1+\exp(-0.5)} = \frac{\exp(-6 + 0.05 * 40 + 3.5)}{1+\exp(-6 + 0.05 * 40 + 3.5)} = \frac{\exp(-0.5)}{1+\exp(-0.5)} = 0.38$$


(b) How many hours would the student in part (a) need to study to have a 50 % chance of getting an A in the class?

***sol.n:***

The log odds is
$$\log(\frac{P(X)}{1 - P(X)}) = \hat\beta_0 + \hat\beta_1X_1 + \hat\beta_2X_2$$;
Then, we use the data from the former question,

$$\log(\frac{0.5}{1 - 0.5}) = -6 + 0.05X_1 + 3.5$$;

$$X_1 = 50$$;

Therefore, the student in part (a) need to study 50 hours to have a 50 % chance of getting an A in the class.


4.8. Suppose that we take a data set, divide it into equally-sized training and test sets, and then try out two different classification procedures.

First we use logistic regression and get an error rate of 20 % on the training data and 30 % on the test data. Next we use 1-nearest neighbors (i.e. K = 1) and get an average error rate (averaged over both test and training data sets) of 18%. Based on these results, which method should we prefer to use for classification of new observations? Why?

***sol.n:***
We should prefer to use for classification of new observations by logistic regression method.

Since when k = 1, we have a training error rate of 0%. However, the average error rate is 18%, which implies the test error rate is 36%. The test error rate of KNN is greater than the test error rate of logistic regression in case of 36% >30%. Therefore, it is better to choose the logistic regression methods because of the lower test error rate.



4.9. This problem has to do with odds.

(a) On average, what fraction of people with an odds of 0.37 of defaulting on their credit card payment will in fact default?

***sol.n:***
The odds ratio is $$\frac{P(X)}{1 - P(X)} = 0.37$$;

Then, $$P(X) = \frac{0.37}{1+ 0.37} = 0.27$$;

Therefore, On average, we have a fraction of 27% people with an odds of 0.37 of defaulting on their credit card payment.



(b) Suppose that an individual has a 16% chance of defaulting on her credit card payment. What are the odds that she will default?

***sol.n:***
Since we have already known P(X) = 0.16;

The odds ratio is $$\frac{P(X)}{1 - P(X)} = \frac{0.16}{1 - 0.16} = 0.19$$;

Therefore, she will default the the odds 0.19.




4.13. This question should be answered using the Weekly data set, which is part of the ISLR2 package. This data is similar in nature to the Smarket data from this chapter’s lab, except that it contains 1, 089 weekly returns for 21 years, from the beginning of 1990 to the end of 2010.

(a) Produce some numerical and graphical summaries of the Weekly data. Do there appear to be any patterns?

```{r}
attach(Weekly)

# look at the feature of data set
summary(Weekly)

# look at the correlation of each variables
cor(Weekly[,-9])
pairs(Weekly)
```
Almost all the correlation is below 0.1.

Additionally, the correlation of "year" and "Volume" is higher than any other correlation values, even higher than 0.7, which illustrates that there exists high relation between "year" and "Volume".


(b) Use the full data set to perform a logistic regression with Direction as the response and the five lag variables plus Volume as predictors. Use the summary function to print the results. Do any of the predictors appear to be statistically significant? If so, which ones?

```{r}
fit.glm <- glm(Direction ~ .-Today - Year, data = Weekly, family = binomial)
summary(fit.glm)
```

When we look at the estimate coefficients, only "Lag2" is significant with p-value which is below 0.05.


(c) Compute the confusion matrix and overall fraction of correct predictions. Explain what the confusion matrix is telling you about the types of mistakes made by logistic regression.

```{r}
probs <- predict(fit.glm, type = "response")

# set down and up
pred.glm <- rep("Down", length(probs))
pred.glm[probs > 0.5] <- "Up"
table(pred.glm, Direction)
```

The overall fraction of correct predictions is (54 + 557)/1089 = 0.5611. 


(d) Now fit the logistic regression model using a training data period from 1990 to 2008, with Lag2 as the only predictor. Compute the confusion matrix and the overall fraction of correct predictions for the held out data (that is, the data from 2009 and 2010).

```{r}
# from 1990 to 2008
train <- (Year < 2009)

Weekly.test <- Weekly[!train, ]
Direction.test <- Direction[!train]

fit.glm2 <- glm(Direction ~ Lag2, data = Weekly, family = binomial, subset = train)
summary(fit.glm2)

probs2 <- predict(fit.glm2, Weekly.test, type = "response")

# set down and up
pred.glm2 <- rep("Down", length(probs2))
pred.glm2[probs2 > 0.5] <- "Up"
table(pred.glm2, Direction.test)
```

The overall fraction of correct predictions is (9 + 56)/(9 + 5 + 34 + 56)= 0.625.


(e) Repeat (d) using LDA.

```{r}
fit.lda <- lda(Direction ~ Lag2, data = Weekly, subset = train)
fit.lda

pred.lda <- predict(fit.lda, Weekly.test)
table(pred.lda$class, Direction.test)
```

The overall fraction of correct predictions is (9 + 56)/(9 + 5 + 34 + 56)= 0.625.


(f) Repeat (d) using QDA.

```{r}
fit.qda <- qda(Direction ~ Lag2, data = Weekly, subset = train)
fit.qda

pred.qda <- predict(fit.qda, Weekly.test)
table(pred.qda$class, Direction.test)
```

The overall fraction of correct predictions is (0 + 61)/(0 + 0 + 43 + 61)= 0.5865.


(g) Repeat (d) using KNN with K = 1.

```{r}
train.X <- as.matrix(Lag2[train])
test.X <- as.matrix(Lag2[!train])
train.Direction <- Direction[train]

set.seed(679)
pred.knn <- knn(train.X, test.X, train.Direction, k = 1)
table(pred.knn, Direction.test)
```

The overall fraction of correct predictions is (21 + 30)/(21 + 30 + 22 + 31)= 0.4904.


(h) Repeat (d) using naive Bayes.

```{r}
fit.nb <- naiveBayes(Direction ~ Lag2, data = Weekly, subset = train)
fit.nb

pred.nb <- predict(fit.nb, Weekly.test)
table(pred.nb, Direction.test)
```
The overall fraction of correct predictions is (0 + 61)/(0 + 0 + 43 + 61)= 0.5865.


(i) Which of these methods appears to provide the best results on this data?

***sol.n:***
If we compare the test error rates, we see that logistic regression and LDA have the minimum error rates, followed by QDA, KNN and naive Bayes.



(j) Experiment with different combinations of predictors, including possible transformations and interactions, for each of the methods. Report the variables, method, and associated confusion matrix that appears to provide the best results on the held out data. Note that you should also experiment with values for K in the KNN classifier.

```{r}
# Logistic regression with Lag2 + Lag1
fit.glm3 <- glm(Direction ~ Lag2 + Lag1, data = Weekly, family = binomial, subset = train)
probs3 <- predict(fit.glm3, Weekly.test, type = "response")
pred.glm3 <- rep("Down", length(probs3))
pred.glm3[probs3 > 0.5] = "Up"
table(pred.glm3, Direction.test)

mean(pred.glm3 == Direction.test)

# LDA with Lag2 + Lag1
fit.lda2 <- lda(Direction ~ Lag2 + Lag1, data = Weekly, subset = train)
pred.lda2 <- predict(fit.lda2, Weekly.test)
table(pred.lda2$class, Direction.test)

mean(pred.lda2$class == Direction.test)

# QDA with Lag2 + Lag1
fit.qda2 <- qda(Direction ~ Lag2 + Lag1, data = Weekly, subset = train)
pred.qda2 <- predict(fit.qda2, Weekly.test)
table(pred.qda2$class, Direction.test)  

mean(pred.qda2$class == Direction.test)

# naive Bayes with Lag2 + Lag1
fit.nb2 <- naiveBayes(Direction ~ Lag2 + Lag1, data = Weekly, subset = train)
pred.nb2 <- predict(fit.nb2, Weekly.test)
table(pred.nb2, Direction.test)

mean(pred.nb2 == Direction.test)

# KNN k =10
pred.knn2 <- knn(train.X, test.X, train.Direction, k = 10)
table(pred.knn2, Direction.test)

mean(pred.knn2 == Direction.test)

# KNN k =100
pred.knn3 <- knn(train.X, test.X, train.Direction, k = 100)
table(pred.knn3, Direction.test)

mean(pred.knn3 == Direction.test)
```

According to all the results, the KNN with k = 10 have the best performance in terms of test correct rates.




4.14. In this problem, you will develop a model to predict whether a given car gets high or low gas mileage based on the Auto data set.

(a) Create a binary variable, mpg01, that contains a 1 if mpg contains a value above its median, and a 0 if mpg contains a value below its median. You can compute the median using the median() function. Note you may find it helpful to use the data.frame() function to create a single data set containing both mpg01 and the other Auto variables.

```{r comment=FALSE, message=FALSE, warning=FALSE}
# create a binary variable "mpg01"
attach(Auto)
Auto$mpg01 = with(ifelse(mpg>median(mpg), 1, 0), data=Auto)
```

(b) Explore the data graphically in order to investigate the association between mpg01 and the other features. Which of the other features seem most likely to be useful in predicting mpg01? Scatterplots and boxplots may be useful tools to answer this question. Describe your findings.

```{r}
# correlation
cor(Auto[, -9])

# correlation plot (scatter plot)
pairs(Auto)

# boxplots
# cylinders
ggplot(Auto, aes(x=as.factor(mpg01), y=cylinders)) + 
  geom_boxplot(fill="slateblue", alpha=0.2) + 
  xlab("mpg") +
  ggtitle("cylinders vs. mpg") 
  
# displacement
ggplot(Auto, aes(x=as.factor(mpg01), y=displacement)) + 
  geom_boxplot(fill="slateblue", alpha=0.2) + 
  xlab("mpg") +
  ggtitle("displacement vs. mpg")

# horsepower
ggplot(Auto, aes(x=as.factor(mpg01), y=horsepower)) + 
  geom_boxplot(fill="slateblue", alpha=0.2) + 
  xlab("mpg") +
  ggtitle("horsepower vs. mpg")

# weight 
ggplot(Auto, aes(x=as.factor(mpg01), y=weight)) + 
  geom_boxplot(fill="slateblue", alpha=0.2) + 
  xlab("mpg") +
  ggtitle("weight vs. mpg")

#acceleration    
ggplot(Auto, aes(x=as.factor(mpg01), y=acceleration)) + 
  geom_boxplot(fill="slateblue", alpha=0.2) + 
  xlab("mpg") +
  ggtitle("acceleration vs. mpg")

#year  
ggplot(Auto, aes(x=as.factor(mpg01), y=year)) + 
  geom_boxplot(fill="slateblue", alpha=0.2) + 
  xlab("mpg") +
  ggtitle("year vs. mpg")
```


(c) Split the data into a training set and a test set.

```{r}
# Split the data randomly
# Training and Test data
set.seed(679)
sample_data = sample.split(Auto$mpg, SplitRatio = 0.70)
Auto.train = subset(Auto, sample_data==TRUE) 
Auto.test = subset(Auto, sample_data==FALSE)
```

(d) Perform LDA on the training data in order to predict mpg01 using the variables that seemed most associated with mpg01 in (b). What is the test error of the model obtained?

```{r}
auto.lda <- lda(mpg01 ~ cylinders + weight + displacement + horsepower, data = Auto, subset = train)
auto.lda

pred.lda3 <- predict(auto.lda, Auto.test)
table(pred.lda3$class, Auto.test$mpg01)

1 - mean(pred.lda3$class == Auto.test$mpg01)
```

The test error of the model obtained by using LAD is 11.54%.


(e) Perform QDA on the training data in order to predict mpg01 using the variables that seemed most associated with mpg01 in (b). What is the test error of the model obtained?

```{r}
auto.qda <- qda(mpg01 ~ cylinders + weight + displacement + horsepower, data = Auto, subset = train)
auto.qda

pred.qda3 <- predict(auto.qda, Auto.test)
table(pred.qda3$class, Auto.test$mpg01)

1 - mean(pred.qda3$class == Auto.test$mpg01)
```

The test error of the model obtained by using QDA is 10.58%.


(f) Perform logistic regression on the training data in order to predict mpg01 using the variables that seemed most associated with mpg01 in (b). What is the test error of the model obtained?

```{r}
auto.glm <- glm(mpg01 ~ cylinders + weight + displacement + horsepower, data = Auto.train, family = binomial)
auto.probs <- predict(auto.glm, Auto.test, type = "response")
pred <- rep("0", length(auto.probs))
pred[auto.probs > 0.5] = "1"
table(pred, Auto.test$mpg01)

1 - mean(pred == Auto.test$mpg01)
```

The test error of the model obtained by using the logistic regression is 9.615%.


(g) Perform naive Bayes on the training data in order to predict mpg01 using the variables that seemed most associated with mpg01 in (b). What is the test error of the model obtained?

```{r}
auto.nb <- naiveBayes(mpg01 ~ cylinders + weight + displacement + horsepower, data = Auto.train)
pred.nb3 <- predict(auto.nb, Auto.test)
table(pred.nb3, Auto.test$mpg01)

1 - mean(pred.nb3 == Auto.test$mpg01)
```

The test error of the model obtained by using naive Bayes is 10.58%.


(h) Perform KNN on the training data, with several values of K, in order to predict mpg01. Use only the variables that seemed most associated with mpg01 in (b). What test errors do you obtain? Which value of K seems to perform the best on this data set?

```{r}
# Train, Test and response matrices.
train.matrix = data.matrix(Auto.train[,c("cylinders","displacement","weight","horsepower")])
test.matrix = data.matrix(Auto.test[,c("cylinders","displacement","weight","horsepower")])
train.y = data.matrix(Auto.train$mpg01)
test.y = data.matrix(Auto.test$mpg01)

# K=1 and predictions
auto.knn1 = knn(train.matrix, test.matrix, train.y, k=1)
table(auto.knn1, test.y)

1 - mean(auto.knn1 == Auto.test$mpg01)
```

The test error of the model obtained by using KNN(k = 1) is 16.35%.

```{r}
# K=10 and predictions
auto.knn2 = knn(train.matrix, test.matrix, train.y, k=10)
table(auto.knn2, test.y)

1 - mean(auto.knn2 == Auto.test$mpg01)
```

The test error of the model obtained by using KNN(k = 10) is 8.654%.

```{r}
# K=100 and predictions
auto.knn3 = knn(train.matrix, test.matrix, train.y, k=100)
table(auto.knn3, test.y)

1 - mean(auto.knn3 == Auto.test$mpg01)
```

The test error of the model obtained by using KNN(k = 100) is 13.46%.
When K is 10,  this data set performs best.



4.15. This problem involves writing functions.
(a) Write a function, Power(), that prints out the result of raising 2 to the 3rd power. In other words, your function should compute 23 and print out the results.
*Hint: Recall that x^a raises x to the power a. Use the print() function to output the result.*

```{r}
Power <- function(a,b){a^b}
Power(2,3)
```

(b) Create a new function, Power2(), that allows you to pass any two numbers, x and a, and prints out the value of x^a. You can do this by beginning your function with the line.
Power2 <- function(x, a) {
You should be able to call your function by entering, for instance,
Power2(3, 8)
on the command line. This should output the value of 38, namely, 6, 561.

```{r}
Power2 <- function(x, a) print(x^a)
```


(c) Using the Power2() function that you just wrote, compute 10^3, 81^7, and 131^3.

```{r}
Power2(10, 3)
Power2(81,7)
Power2(131,3)
```

(d) Now create a new function, Power3(), that actually returns the result x^a as an R object, rather than simply printing it to the screen. That is, if you store the value x^a in an object called result within your function, then you can simply return() this result, using the following line:
return(result)
The line above should be the last line in your function, before
the } symbol.


```{r}
Power3 <- function(x, a) x^a
```


(e) Now using the Power3() function, create a plot of f(x) = x2. The x-axis should display a range of integers from 1 to 10, and the y-axis should display x2. Label the axes appropriately, and use an appropriate title for the figure. Consider displaying either the x-axis, the y-axis, or both on the log-scale. You can do this by using log = "x", log = "y", or log = "xy" as arguments to the plot() function.

```{r}
x <- seq(from = 1, to = 10, length.out = 100)

plot(x, Power3(x, 2), type = 'l', 
     ylab = expression(x^2),
     main = expression('f(x) = '*x^2*' in [0,10]'))
```


(f) Create a function, PlotPower(), that allows you to create a plot of x against x^a for a fixed a and for a range of values of x. For instance, if you call

```{r}
# PlotPower(1:10, 3)
```

then a plot should be created with an x-axis taking on values 1,2,...,10, and a y-axis taking on values 13,23,...,103.


```{r}
PlotPower <- function(x, a) {
  
    plot(x, Power3(x, a), type = 'l', xlab = 'x', ylab = 'f(x)',
         main = paste('Plot of x to the power of', a))
}

PlotPower(1:10, 3)
```



4.16. Using the Boston data set, fit classification models in order to predict whether a given census tract has a crime rate above or below the median. Explore logistic regression, LDA, naive Bayes, and KNN models using various subsets of the predictors. Describe your findings.
*Hint: You will have to create the response variable yourself, using the variables that are contained in the Boston data set.*

```{r}
# Boston data set
library(MASS)
attach(Boston)
```

```{r}
# fit classification models in order to predict whether a given census tract has a crime rate above or below the median.
# add 1 to a new column if crim > median, o otherwise.

Boston$crim01 = with(ifelse(crim>median(crim), 1, 0), data=Boston)

# check the correlation
cor(Boston)

# training and test set
set.seed(679)
Boston.sample = sample.split(Boston$crim01, SplitRatio = 0.80)
Boston.train = subset(Boston, Boston.sample==TRUE) 
Boston.test = subset(Boston, Boston.sample==FALSE)
```


```{r}
# Logistic regression
boston.glm <- glm(crim01 ~ indus + nox + age + rad + tax, data = Boston.train, family = binomial)
summary(boston.glm)

boston.probs <- predict(boston.glm, Boston.test, type = "response")
boston.pred <- rep("0", length(boston.probs))
boston.pred[boston.probs > 0.5] = "1"
table(boston.pred, Boston.test$crim01)

1 - mean(boston.pred == Boston.test$crim01)
```


```{r}
# LDA
boston.lda <- lda(crim01 ~ indus + nox + age + rad + tax, data = Boston.train)
boston.lda

pred.blda <- predict(boston.lda, Boston.test)
table(pred.blda$class, Boston.test$crim01)

1 - mean(pred.blda$class == Boston.test$crim01)
```


```{r}
# Naive Bayes
boston.nb <- naiveBayes(crim01 ~ indus + nox + age + rad + tax, data = Boston.train)
pred.bnb <- predict(boston.nb, Boston.test)
table(pred.bnb, Boston.test$crim01)

1 - mean(pred.bnb == Boston.test$crim01)
```


```{r}
# KNN models
# "indus","nox","age","rad","tax"
# indus,nox,age,dis,rad,tax
train.matrix = data.matrix(Boston.train[,c("indus","nox","age","rad","tax")])
test.matrix = data.matrix(Boston.test[,c("indus","nox","age","rad","tax")])
train.y = data.matrix(Boston.train$crim01)
test.y = data.matrix(Boston.test$crim01)

# K=1 and predictions
boston.knn1 = knn(train.matrix, test.matrix, train.y, k=1)
table(boston.knn1, test.y)

1 - mean(boston.knn1 == Boston.test$crim01)



# K=5 and predictions
boston.knn2 = knn(train.matrix, test.matrix, train.y, k=5)
table(boston.knn2, test.y)

1 - mean(boston.knn2 == Boston.test$crim01)
```
The result shows that KNN(k = 1) performs the best among these methods with an test error rate around 10%.
